\textbf{The statistical analysis of non-linear longitudinal data in biomedical research using generalized additive models}

### 2021-01-20 Manuscripts/Manuscript_AM_v2.Rmd
Manuscript and figures updated

---
title: '**The statistical analysis of non-linear longitudinal data in biomedical research using generalized additive models**'
subtitle: _Beyond repeated measures ANOVA and Linear Mixed Models_
author:
- Ariel Mundo^[Department of Biomedical Engineering, University of Arkansas, Fayetteville]
- Timothy J. Muldoon^[Department of Biomedical Engineering, University of Arkansas,
  Fayetteville]
- John R. Tipton^[Department of Mathematical Sciences, University of Arkansas, Fayetteville]
output:
  rticles::jss_article
  bookdown::html_document2: default
  bookdown::word_document2: default
  bookdown::pdf_document2:
    keep_tex: yes
    fig_caption: yes
    extra_dependencies: ["subfig","breqn"]
  link-citations: yes
  css: style.css
  '': default
csl: Elsevier.csl
bibliography: refs.bib
---

@@

plot_example <- function(sim_dat) {
  library(patchwork)
  ## Plot the simulated data
  p1 <- sim_dat$dat %>%
    ggplot(aes(x = time, y = y, group = treatment, color = treatment)) +
    geom_point() +
    geom_line(aes(x = time, y = mu, color = treatment)) +
    theme_bw(base_size = 12) +
    ggtitle("Simulated data with true response function")+
    scale_color_aaas()

  p2 <- sim_dat$dat %>%
    ggplot(aes(x = time, y = y, group = subject, color = treatment)) +
    geom_line(aes(size = "Subjects")) +
    # facet_wrap(~ treatment) +
    geom_line(aes(x = time, y = mu, color = treatment, size = "Simulated Truth"), lty = 1) +
    scale_size_manual(name = "Type", values=c("Subjects" = 0.5, "Simulated Truth" = 3)) +
    ggtitle("Simulated data\nIndividual responses with population mean") +
    theme_bw(base_size = 12)+
    scale_color_aaas()

   p3 <- sim_dat$dat %>%
    ggplot(aes(x = time, y = errors, group = subject, color = treatment)) +
    geom_line() +
    # facet_wrap(~ treatment) +
    ggtitle("Simulated errors") +
    theme_bw(base_size = 12)+
    scale_color_aaas()

  p4 <- ggplot(sim_dat$dat, aes(x = time, y = y, color = treatment)) +
    geom_point() +
    geom_line(aes(y = predict(sim_dat$fit_lme), group = subject, size = "Subjects")) +
    geom_line(data = sim_dat$pred_dat, aes(y = predict(sim_dat$fit_lme, level = 0, newdata = sim_dat$pred_dat), size = "Population")) +
    scale_size_manual(name = "Predictions", values=c("Subjects" = 0.5, "Population" = 3)) +
    theme_bw(base_size = 12) +
    ggtitle("Fitted Model")+
    scale_color_aaas()

@@

```{r, linear-model, fig.width=16, fig.height=9, out.width="100%", echo=FALSE,message=FALSE,fig.show='hold',fig.cap="Simulated linear responses from two groups with a rm-ANOVA model fitted. Top row, simulated data, lines represent mean response. Bottom row, fitted model, thick lines represent predicted mean response"}
# linear response, correlated errors (subject effect)
txt<-20
plot_example(example(fun_type = "linear", error_type = "correlated")) +
  plot_annotation(
  title = 'Simulated with linear function and correlated errors',
    theme = theme(plot.title = element_text(size = txt),text=element_text(size=txt))
)
```

@@

```{r quadratic-response, fig.width=16, fig.height=9, out.width="100%",echo=FALSE,fig.cap='Simulated quadratic responses from two groups with an rm-ANOVA model fitted. Top row, simulated data, lines represent mean response. Bottom row, fitted model, thick lines represent predicted mean response', message=FALSE}
# quadratic response, correlated errors (subject effect)
library(ggsci)
plot_example(example(fun_type = "linear", error_type = "correlated")) +
  plot_annotation(
  title = 'Simulated with linear function and correlated errors',
    theme = theme(plot.title = element_text(size = txt),text=element_text(size=txt))
)
  # theme(plot.title = element_text(size = 20))
```

>>>>

---
title: '**The statistical analysis of non-linear longitudinal data in biomedical research using generalized additive models**'
subtitle: _Beyond repeated measures ANOVA and Linear Mixed Models_
author:
- Ariel Mundo^[Department of Biomedical Engineering, University of Arkansas, Fayetteville]
- Timothy J. Muldoon^[Department of Biomedical Engineering, University of Arkansas,
  Fayetteville]
- John R. Tipton^[Department of Mathematical Sciences, University of Arkansas, Fayetteville]
output:
  bookdown::pdf_document2:
    keep_tex: yes
    fig_caption: yes
    extra_dependencies: ["subfig","breqn"]
  bookdown::word_document2: default
  bookdown::html_document2: default
  link-citations: yes
  css: style.css
  '': default
csl: Elsevier.csl
bibliography: refs.bib
---

@@

plot_example <- function(sim_dat) {
  library(patchwork)
  ## Plot the simulated data
  p1 <- sim_dat$dat %>%
    ggplot(aes(x = time, y = y, group = treatment, color = treatment)) +
    geom_point() +
    geom_line(aes(x = time, y = mu, color = treatment)) +
    theme_classic() +
    ggtitle("Simulated data with true response function")+
    theme(plot.title = element_text(size = 30, 
                                  face = "bold"),
        text=element_text(size=30))+
    scale_color_aaas()

  p2 <- sim_dat$dat %>%
    ggplot(aes(x = time, y = y, group = subject, color = treatment)) +
    geom_line(aes(size = "Subjects")) +
    # facet_wrap(~ treatment) +
    geom_line(aes(x = time, y = mu, color = treatment, size = "Simulated Truth"), lty = 1) +
    scale_size_manual(name = "Type", values=c("Subjects" = 0.5, "Simulated Truth" = 3)) +
    ggtitle("Simulated data\nIndividual responses with population mean") +
    theme(plot.title = element_text(size = 30, 
                                  face = "bold"),
        text=element_text(size=30))+
    theme_classic()+
    scale_color_aaas()

   p3 <- sim_dat$dat %>%
    ggplot(aes(x = time, y = errors, group = subject, color = treatment)) +
    geom_line() +
    # facet_wrap(~ treatment) +
    ggtitle("Simulated errors") +
     theme(plot.title = element_text(size = 30, 
                                  face = "bold"),
        text=element_text(size=30))+
    theme_classic()+
    scale_color_aaas()

  p4 <- ggplot(sim_dat$dat, aes(x = time, y = y, color = treatment)) +
    geom_point() +
    geom_line(aes(y = predict(sim_dat$fit_lme), group = subject, size = "Subjects")) +
    geom_line(data = sim_dat$pred_dat, aes(y = predict(sim_dat$fit_lme, level = 0, newdata = sim_dat$pred_dat), size = "Population")) +
    scale_size_manual(name = "Predictions", values=c("Subjects" = 0.5, "Population" = 3)) +
    theme_classic() +
    ggtitle("Fitted Model")+
    theme(plot.title = element_text(size = 30, 
                                  face = "bold"),
        text=element_text(size=30))+
    scale_color_aaas()

@@

```{r, linear-model, fig.width=16, fig.height=12,  echo=FALSE,message=FALSE,fig.show='hold',fig.cap="Simulated linear responses from two groups with a rm-ANOVA model fitted. Top row, simulated data, lines represent mean response. Bottom row, fitted model, thick lines represent predicted mean response"}
# linear response, correlated errors (subject effect)
txt<-18
plot_example(example(fun_type = "linear", error_type = "correlated")) +
  plot_annotation(tag_levels='A',
                  title = 'Simulated with linear function and correlated errors') &
   theme(plot.title = element_text(size = 20),text=element_text(size=txt)
)
```

@@

```{r quadratic-response, fig.width=16, fig.height=16,echo=FALSE,fig.cap='Simulated quadratic responses from two groups with an rm-ANOVA model fitted. Top row, simulated data, lines represent mean response. Bottom row, fitted model, thick lines represent predicted mean response', message=FALSE}
# quadratic response, correlated errors (subject effect)
txt<-18
plot_example(example(fun_type = "linear", error_type = "correlated")) +
  plot_annotation(
    tag_levels='A',
  title = 'Simulated with linear function and correlated errors') &
   theme(plot.title = element_text(size = 20),text=element_text(size=txt)
)
  # theme(plot.title = element_text(size = 20))
```

### 2021-01-20 Manuscripts/Manuscript_AM_v2.tex
Manuscript and figures updated

\title{\textbf{The statistical analysis of non-linear longitudinal data in biomedical research using generalized additive models}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{\emph{Beyond repeated measures ANOVA and Linear Mixed Models}}
\author{true \and true \and Ariel Mundo\footnote{Department of Biomedical Engineering, University of Arkansas, Fayetteville} \and Timothy J. Muldoon\footnote{Department of Biomedical Engineering, University of Arkansas, Fayetteville} \and John R. Tipton\footnote{Department of Mathematical Sciences, University of Arkansas, Fayetteville}}
\date{}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{Manuscript_AM_v2_files/figure-latex/linear-model-1} \caption{Simulated linear responses from two groups with a rm-ANOVA model fitted. Top row, simulated data, lines represent mean response. Bottom row, fitted model, thick lines represent predicted mean response}\label{fig:linear-model}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{Manuscript_AM_v2_files/figure-latex/quadratic-response-1} \caption{Simulated quadratic responses from two groups with an rm-ANOVA model fitted. Top row, simulated data, lines represent mean response. Bottom row, fitted model, thick lines represent predicted mean response}\label{fig:quadratic-response}
\end{figure}

>>>>

\title{\textbf{The statistical analysis of non-linear longitudinal data in biomedical research using generalized additive models}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{\emph{Beyond repeated measures ANOVA and Linear Mixed Models}}
\author{Ariel Mundo\footnote{Department of Biomedical Engineering, University of Arkansas, Fayetteville} \and Timothy J. Muldoon\footnote{Department of Biomedical Engineering, University of Arkansas, Fayetteville} \and John R. Tipton\footnote{Department of Mathematical Sciences, University of Arkansas, Fayetteville}}
\date{}

@@

\begin{figure}
\centering
\includegraphics{Manuscript_AM_v2_files/figure-latex/linear-model-1.pdf}
\caption{\label{fig:linear-model}Simulated linear responses from two groups with a rm-ANOVA model fitted. Top row, simulated data, lines represent mean response. Bottom row, fitted model, thick lines represent predicted mean response}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{Manuscript_AM_v2_files/figure-latex/quadratic-response-1.pdf}
\caption{\label{fig:quadratic-response}Simulated quadratic responses from two groups with an rm-ANOVA model fitted. Top row, simulated data, lines represent mean response. Bottom row, fitted model, thick lines represent predicted mean response}
\end{figure}
